Efficient Machine Learning in a Memristor Network

### 2025-06-04 paper/paper.tex
Updated graphs in paper

\begin{abstract}
Inspired by the brain's energy-efficient processing, we propose a physical implementation of artificial neural networks to surpass digital systems in speed and energy efficiency. We consider a network with nodes connected by conical microfluidic memristor channels filled with aqueous electrolyte [1]. These memristors exhibit tunable conductance, responsive to electrical, chemical, and mechanical stimuli, as well as channel geometry. Based on this adaptability, we develop a hybrid training algorithm that adjusts either channel geometry or external stimuli as weights. Physical relaxation replaces the feedforward pass, while a combined physical-numerical gradient descent drives the backward pass. This approach enables tasks such as input-output voltage mapping and linear regression.
\end{abstract}
\maketitle

@@

Modern computing algortihms are able to efficiently perform complicated nuemrical and symbolic operations, but their sequential structure limits their performance on more perceptual tasks. Inspired by the architecture and the functioning of the human brain, that performs tasks via distributed and parallel fashion, artificial neural network (ANN) algorithms outreached traditional computing achieving remarkable success in challenging tasks, such as visual and audio pattern recognition~\cite{pattern_recognition,review_image_rec, speech_rec} to natural language processing and became central to many technological advancements in the field of artificial intelligence\cite{ai_tech1}.

@@

The PNN we consider is an electrical network made of nodes connected by resistive elements. As resistive elements, we consider conical microfluidic channels, schematically represented in \cref{fig:memristor_scheme}. 
\begin{figure}[h]
    \centering
    \includegraphics[width=\columnwidth]{plots/memristor/memristor_scheme.pdf}
    \caption{A schematic representation of the conical fluidic memristor. A cone of lenght $L$, base radius $R_b$ and tip radius $R_t$ with charged surface connnects two bulk reservoirs of a 1:1 electrolyte in an acqueous solution. The bulk density in the base $\rho_b$ differs from the bulk density in the tip $\rho_t$. At the far side of the memristor, an electrical potential and pressure drop are applied. Overall, this results in a chemical, electrical and mechanical sensistive system.}\label{fig:memristor_scheme}
\end{figure} 
An azimuthally symmetric conical channel connects two reservoirs that contain an incompressible aqueous 1:1 electrolyte with viscosity $\eta=1.01mPa s$ containing ions with diffusion coefficients $D = 1.75\mu m^2 ms^{-1}$ and charge $\pm e$, with $e$ the proton charge. The cone has a base radius of $R_b = 200$nm, a tip radius of $R_t=50$nm and has length $L = 10\mu$m, which sets it in the long channel limit, where entrance and exit effects can be neglected. The walls of the channel are homogeneously charged and generate a surface potential $\psi_0$ that attracts the opposite-sign ions present in the solution due to the dissociation of the electrolyte. The electric potential decreases exponentially away from the surface due to the electrostatic screening of the attracted ions: an electric double layer (EDL) forms on the charged surfaces.

At the far side of both reservoirs, electrical potential, pressure and ion concentration are imposed. As a result, the memristor feels a potential drop $\Delta V = V_b - V_t$ defined as the difference between the potential in the base and in the tip. The ions within the EDL region generate an electro-osmotic fluid flow $Q_V = (\pi R_b R_t\epsilon\psi_0/L\eta)\Delta V$. Meanwhile, the pressure drop $\Delta P = P_b - P_t$, given by the difference between the base and the tip pressure, induces a Poiseuille-like flow $Q_P = (\pi R_b^3 R_t^3/8L\eta\langle R^2\rangle)\Delta P$. Finally, the ions experience a concentration gradient $\Delta \rho = \rho_b - \rho_t$, defined by the difference between the ion concentration in the reservoir connected to the base of the channel, $\rho_b$, and the concentration in the reservoir connected to the tip, $\rho_t$. Consequently, a diffusive fluid flow arises. However, we consider only the pressure and potential contributions to the total fluid flow, $Q = Q_P + Q_V$, assuming the diffusive flow to be negligible.  The resulting fluid flow generates non-trivial ion concentration profiles which, in turn, directly infulence the channel's conductance $g$. 
%It has already been shown that, for $\Delta P = 0$, the channel shows current rectification, thus working as a diode \cite{}, due to concentration polarization.

@@

When the external stimuli vary dynamically in time, they generate ion accumulation and deplition in the channel, changing consequently the conductivity. We assume the dynamics of the time-dependent conductance $g(t)$ as a single exponential decay towards the steady-state conductance $g_{\infty}$ with a diffusion-like timescale $\tau = L^2/12D$, as in Ref. \cite{Rene_IonicNeuromorphic}. The equation for the conductance reads
\begin{equation}
\frac{\partial g\brac{\boldsymbol{\Delta}(t), t}}{\partial t} = \frac{g_{\infty}\brac{\boldsymbol{\Delta}(t)}-g\brac{\boldsymbol{\Delta}(t), t}}{\tau}
\label{eq:eqofmotconductance}
\end{equation}
with $\boldsymbol{\Delta}(t) = \brac{\Delta V\brac{t},\Delta P\brac{t}, \Delta \rho\brac{t}}$.
As in Ref. \cite{Rene_IonicNeuromorphic}, we consider the time scale relative to voltage-driven ion accumulation or depletion in the channel. We justify this choice by stating that pressure-driven dynamics is faster and recalling that we neglect chemically-driven transport. Due to the fact that the conductance depends on the history of the signals applied, this system is called memristor. Aside from the direct dependence of the conductance on external stimuli, there is also an implicit dependence on geometrical properties, such as length $L$ and base radius $R_b$, that we address in this work.

The PNN is composed of $\N$ nodes, an example is shown in \cref{fig:memristor_network}. 
\begin{figure}[h]
    \centering
    \includegraphics[width=0.8\columnwidth]{plots/memristor/memristor_network.pdf}
    \caption{An example of a network of five nodes connected by six memristors. The nodes are divided in input nodes (green circles), hidden nodes (gray nodes) and output nodes (purple circles) each of them is characterized by a triplet of physical quantities, representing the potential, pressure, and density. Th resulting stimuli on the memristor is given by the difference between base and tip signal, it thus depends on the oreintation of the memristor. In the case of the memristor that connects the first and the second node, for example, the exteranl stimuli are $\Delta V = V_2 - V_1$, $\Delta P = P_2 - P_1$ and $\Delta \rho = \rho_2 - \rho_1$. The zig-zag lines directed towards the input nodes indicate where the electrical signal is applied, while the zig-zag lines pointing outwards from the nodes indicate where the electrical signal is measured.}
    \label{fig:memristor_network}
\end{figure} 
Each node $\n\in[1,\N]$ is characterized by a value of electrical potential $V_{\n}$, pressure $P_{\n}$ and bulk density $\rho_{\n}$. Even though each node in the network is characterized by three physical properties, only electrical current flows through the network. This current propagates via the memristive elements, enabling modifications to the electric potential at other nodes. In contrast, pressure and density are externally imposed at each node and do not propagate through the network. The set of nodes is divided in thee sets. A node belongs to the set of \textit{input nodes} if it is used to imput current in the circuit, its value of voltage is known and fixed. After the imputs are imposed, the electrical signal travels throught the network and the resulting output potential is read in the nodes belonging to \textit{output nodes}. The remaining nodes, that do not function as input or output, are included in the set of \textit{hidden nodes}.

The nodes in the PNN are connected by $\M$ memristive elements. Consequently, a memristor connecting two nodes experinces potential drop, pressure drop and concentration gradient given by the difference between the node where the base of the memristor is connected to, and the node to which the tip is linked. The memristor is characterized by its conductance state $g_{\m}$ with $\m\in[1,\M]$, consequence of both the geomtery parameters and the external stimuli.

To solve numerically the resulting circuit, it is necessary to simultaneously address the circuit equations and the time evolution equation for the memristors' conductance. 
The circuit equations are handled using the open-source circuit solver \textsc{AHKAB}\cite{}, which computes the electric potentials at the nodes and the currents through the edges discretizing the differential equations in timesteps $\Delta t$. To incorporate the memristive elements, we approximate the time evolution of the conductances using a discretized form of \cref{eq:eqofmotconductance}, employing the Euler-forward method. Specifically, the conductance at a given timestep is calculated from its value at the previous timestep as:
\begin{multline}
g_{\m}\brac{\boldsymbol{\Delta}(t), t} = g_{\m}\brac{\boldsymbol{\Delta}(t), t-\Delta t} + \\
 \frac{g_{\m,\infty}\brac{\boldsymbol{\Delta}(t)-g_{\m}\brac{\boldsymbol{\Delta}, t}}}{\tau}\Delta t.
\end{multline}
In this work, we only apply constant voltage supplies to the input nodes of the circuit. Nonetheless, we notice that solving the dynamical equations is needed due to the dynamical change in the conductance of the memristors that, in turn, results in time-dependent potential drops across them. A more detailed illustration of the algorithm that solves the circuit, along with an analysis of the relaxation to the steady state of the memristor network, is provided in \cref{app:dyn_evol_steady_state}.

@@

We developed a training algorithm that makes use of the physical dependencies of memristor conductivity on both external stimuli and geometric properties to implement a physical neural network. These dependencies allow us to interpret such physical parameters as tunable weights, analogous to standard artificial neural networks, and to encode learning directly into the physical structure of the system. As in traditional ANNs, the goal of the training algorithm is to modify these weights so that the network maps a given set of inputs to a corresponding set of desired outputs. In the case of the memristor network considered here, training aims to produce specific voltage values at designated output nodes, given fixed voltage inputs at the input nodes.
\begin{figure}[!h]
    \centering
    \includegraphics[width=0.8\columnwidth]{plots/training/two_types_weights.pdf}
    \caption{(a) Memristor network with weights defined in the nodes. (b) Memristor network where wegiths characterize each edge.}
    \label{fig:two_types_weights}
\end{figure} 
Crucially, the physical quantities that influence memristor conductance fall into two distinct categories: external stimuli, such as pressure or concentration, that are applied to the nodes, and geometric parameters, such as channel length or base radius, that are defined on the edges. As a result, we distinguish between two types of training strategies: \emph{node-defined weight training}, as illustrated in \cref{fig:two_types_weights}(a), and \emph{edge-defined weight training}, as shown in \cref{fig:two_types_weights}(b).

A modification to any of these weights affects the network's behavior by altering the functional form of the memristor’s steady-state conductance, $g_{\infty}(\boldsymbol{\Delta})$, which depends on the local potential, pressure, and concentration drops across the device. Importantly, the \emph{entire functional shape} of $g_{\infty}$ plays a role in determining the network’s dynamics and ultimate steady state, as illustrated in \cref{app:dyn_evol_steady_state}.

In this work, we focus on training a single property as the weight, while keeping all other parameters fixed as initially defined. For example, we may select an external stimulus, such as the imposed pressure or the density at each node. Alternatively, we can choose the length or base radius of the channel. The training process is carried out by iterating a scheme, illustrated in \cref{fig:training_scheme}, which consists of two phases: a forward pass and a backward pass.
\begin{figure*}[t] 
    \centering
    \includegraphics[width=\textwidth]{plots/training/training_scheme.pdf}
    \caption{Scheme representing the two phases of the training algorithm, they are iteraitvely performed for a selected number of training steps.}
    \label{fig:training_scheme}
\end{figure*}
In the forward phase, the network is fed with specific electric potentials in the input nodes. Following both circuit laws and memristor's dynamics, the network reaches a steady state where electrical potentials can be measured in the output nodes. Clearly, the output of the network depends on the set of weights imposed on the network $\boldsymbol{w}$, regardless them being node or edge defined. The value of the outputs can be used to quantify the status of the network, meaning how much its answer differs from the desired one. Specifically, we define a cost function that computes the difference squared between the voltage in the output nodes and the values of desired voltages
\[
C\left( \boldsymbol{w} \right) = \sum_{i \in \Lambda} \brac{V_i\brac{\boldsymbol{w}} - V_i^D}^2,
\label{eq:cost_func}
\]
where $\Lambda$ represents the set of indeces belonging to the output nodes, $V_i\brac{\boldsymbol{w}}$ the voltage measured on the output nodes and $V^D$ is a vector containing the values of desired output nodes.

The backward phase has the fundamental purpose of finding a set of weights that minimize the cost function. We follow the method of the \textit{steepest descent}, for which the update of the weights from one training step $s$ to the following $s+1$,
\begin{equation}
\boldsymbol{w}^{s+1} = \boldsymbol{w}^{s} - \alpha \boldsymbol{\nabla}C\brac{\boldsymbol{w}},
\label{eq:weight_update}
\end{equation}
results in a decrease in the cost funtion until a minimum is reached, for sufficienlty small values of $\alpha$ \cite{steepest_descent_book}. The gradient in the update rule is computed using a \textit{finite differences method}
\begin{equation}
\boldsymbol{\nabla}C\brac{\boldsymbol{w}} = \frac{C\brac{\boldsymbol{w}+\boldsymbol{\Delta w}}-C\brac{\boldsymbol{w}}}{\Delta w},
\label{eq:gradient_cost_func}
\end{equation}
where $\boldsymbol{\Delta w}$ is a vector with the same number of entries as the weight vector, but every entry has a value $\Delta w$.

In the training algortihm, we first initialize the system with the memristor parameters specified in \cref{sec:memristor_network}, an initial pressure of $1bar$ and concentration of $0.2mM$ at each node. With the knowledge of input and desired output voltage values, it is possible to performe the first forward pass, from which we obtain the cost function $C\brac{\boldsymbol{w}}$, that we store since it is needed in the computation its gradient. Subsequently, we proceed with the backward pass. We compute each entry of the gradient in \cref{eq:gradient_cost_func} by selecting the corresponding weight, a node or an edge parameter, increment its value of $\Delta w$ and applying a forward step to obtain the corresponding entry of $C\brac{\boldsymbol{w}+\boldsymbol{\Delta w}}$. With the stored value of $C\brac{\boldsymbol{w}}$, we are able to compute the gradient in \cref{eq:gradient_cost_func} and use it to calculate the weight update in \cref{eq:weight_update}.

Althought performed numerically in this work, we highlight that the feedforward pass is a physical process in which the system natually relaxes to its steady state. Moreover, also the backward pass we use can be translated to a physical event since involves performing the forward pass a number of time as the number of the weights. Once the gradient of the cost function is determined and stored, only the simple algebric calculations of \cref{eq:weight_update} are left to be computed. We propose this method as a hybrid physical and in-silico training algorithm, as a future experimental approach to the probelm.

@@

We initialize the system with pressures $P_1 =  P_2 = 1 bar$ and ion concetration $\rho_1 = \rho_2 = 0.2 mM$ on each node, length $L_1 = L_2 = 10\mu m$ and base radius  $R_{b1} = R_{b2} = 200 nm$ on each edge. After choosing one of these four physical quantities as training parameter $\boldsymbol{w} = \brac{w_1, w_2}$, we implement the training as descibed in \cref{sec:training} that modifies them, while the other parameters are kept constant. 
In \cref{fig:mse_weights_vd} we show that the cost function normalized to its initial value decreases of many order of magnitude during the training for each of the choosen weight. The values of the choosen set of weights simultaneouly changes from the initial condition, as shown in Figures [], adjusting the behaviour of the network to the desired one. If the values of a choosen type of weight at the end of the training are imposed to the netowork together with the voltage inputs, the network will autonomously realx to a staedy state that coincides with the desired state. To exemplify this point, the inset in Figure shows the physical relaxation of the voltage divider when the trained values of pressure, which correspond to the three values in the last training step, are imposed on the nodes together with the constant inputs voltages. The inset shows that the output voltage $V_2$ autonomously relaxes to the desired value of $V_D=4V$ in a time scale characteristic of the memristor.

\begin{figure}[h]
    \centering
    \includegraphics[width=\columnwidth]{plots/voltage_divider/mse_weights_vd.pdf}
    \caption{(a) Behaviour of the cost function during the training of the voltage divider. Different colors indicate that the training is done with a specific choise of training weights, as indiacted in the legend. (b) Pressures imposed on the three nodes of the voltage divider during training. The inset shows the relaxation to the steady state of the output voltage $V_2$ to the desired value $V_D$ when the pressures imposed are the values obtained in the last training step, and the input nodes are stimulated with fixed voltage. (c) Ion concentration desnities imposed on the three nodes of the voltage divider during training. (d) Length of the memristors during training. (e) Base radius of the memristors during training.}\label{fig:mse_weights_vd}
\end{figure}

In order to investigate the adaptability of the algorithm, we trained the system with a set of three consequent desired voltages $V_D=\{1V, 3V, 4V\}$. In \cref{fig:evolution_targets} we show how the output voltgage $V_2$ subsequently adapts to the desired voltages during the training of the pressures in the nodes, the evolution of which can be seen in \cref{fig:evolution_targets}. It is evident that the system can be retrained at any stage to a new desired objective making it reusable. Any other weight type can be chosen, depending on the necessity.

\begin{figure}[h]
    \centering
    \includegraphics[width=0.8\columnwidth]{plots/voltage_divider/evolution_targets.pdf}
    \caption{(a) The trainiang of a memristor voltage divider to a set of desired voltage outputs $V_D = {1V, 3V, 4V}$, indicated by dashed light purple lines. The output voltage $V_2$, indicated by dark purple dots, adapts to the desired output during the training. The input voltages $V_1$ and $V_3$ are constant at $5V$ and $0V$ respectively. (b) The dots with different shades of gray show the values of pressures, chosen as weights, imposed on the three nodes during training.}\label{fig:evolution_targets}
\end{figure}

@@

To generalize the training objective, we consider a memristor network constructed with a more complex geometry. \cref{fig:mse_general}(a) illustrates a network with three input nodes and two output nodes. The network is trained to produce output voltages of $[3,V, 4,V]$ when provided with input voltages of $[0,V, 2,V, 5,V]$.  Reaching a desired input-output voltage mapping in a generic electric circuit can be useful for variuous electronic purpuses \cite{voltage_controlled_oscillator, onchip_powerdistributed}.

>>>>

\begin{abstract}
Inspired by the brain's energy-efficient processing, we propose a physical implementation of artificial neural networks to surpass digital systems in speed and energy efficiency. We consider a network with nodes connected by conical microfluidic memristor channels filled with aqueous electrolyte [1]. These memristors exhibit tunable conductance, responsive to electrical, chemical, and mechanical stimuli, as well as channel geometry. Based on this adaptability, we develop a hybrid training algorithm that adjusts either channel geometry or external stimuli as weights. Physical relaxation replaces the feedforward pass, while a combined physical-numerical gradient descent drives the backward pass. This approach enables tasks such as input-output voltage mapping and linear regression with bias.
\end{abstract}
\maketitle

@@

Modern computing algorithms are able to efficiently perform complicated numerical and symbolic operations, but their sequential structure limits their performance on more perceptual tasks. Inspired by the architecture and the functioning of the human brain, that performs tasks via distributed and parallel fashion, artificial neural network (ANN) algorithms outreached traditional computing achieving remarkable success in challenging tasks, such as visual and audio pattern recognition~\cite{pattern_recognition,review_image_rec, speech_rec} to natural language processing and became central to many technological advancements in the field of artificial intelligence\cite{ai_tech1}.

@@

The PNN we consider is an electrical network made of nodes connected by resistive elements. As resistive elements, we consider conical microfluidic channels, schematically represented in \cref{fig:memristor_scheme}. 
\begin{figure}[h]
    \centering
    \includegraphics[width=\columnwidth]{plots/memristor/memristor_scheme.pdf}
    \caption{A schematic representation of the conical fluidic memristor. A cone of length $L$, base radius $R_b$ and tip radius $R_t$ with charged surface connects two bulk reservoirs of a 1:1 electrolyte in an aqueous solution. The bulk density in the base $\rho_b$ differs from the bulk density in the tip $\rho_t$. At the far side of the memristor, an electrical potential and pressure drop are applied. Overall, this results in a chemical, electrical and mechanical sensitive system.}\label{fig:memristor_scheme}
\end{figure} 
An azimuthally symmetric conical channel connects two reservoirs that contain an incompressible aqueous 1:1 electrolyte with viscosity $\eta=1.01mPa s$ containing ions with diffusion coefficients $D = 1.75\mu m^2 ms^{-1}$ and charge $\pm e$, with $e$ the proton charge. The cone has a base radius of $R_b = 200$nm, a tip radius of $R_t=50$nm and has length $L = 10\mu$m, which sets it in the long channel limit, where entrance and exit effects can be neglected. The walls of the channel are homogeneously charged and generate a surface potential $\psi_0$ that attracts the opposite-sign ions present in the solution due to the dissociation of the electrolyte. The electric potential decreases exponentially away from the surface due to the electrostatic screening of the attracted ions: an electric double layer (EDL) forms on the charged surfaces.

At the far side of both reservoirs, electrical potential, pressure and ion concentration are imposed. As a result, the memristor feels a potential drop $\Delta V = V_b - V_t$ defined as the difference between the potential in the base and in the tip. The ions within the EDL region generate an electro-osmotic fluid flow $Q_V = (\pi R_b R_t\epsilon\psi_0/L\eta)\Delta V$. Meanwhile, the pressure drop $\Delta P = P_b - P_t$, given by the difference between the base and the tip pressure, induces a Poiseuille-like flow $Q_P = (\pi R_b^3 R_t^3/8L\eta\langle R^2\rangle)\Delta P$. Finally, the ions experience a concentration gradient $\Delta \rho = \rho_b - \rho_t$, defined by the difference between the ion concentration in the reservoir connected to the base of the channel, $\rho_b$, and the concentration in the reservoir connected to the tip, $\rho_t$. Consequently, a diffusive fluid flow arises. However, we consider only the pressure and potential contributions to the total fluid flow, $Q = Q_P + Q_V$, assuming the diffusive flow to be negligible.  The resulting fluid flow generates non-trivial ion concentration profiles which, in turn, directly influence the channel's conductance $g$. 
%It has already been shown that, for $\Delta P = 0$, the channel shows current rectification, thus working as a diode \cite{}, due to concentration polarization.

@@

When the external stimuli vary dynamically in time, they generate ion accumulation and depletion in the channel, changing consequently the conductivity. We assume the dynamics of the time-dependent conductance $g(t)$ as a single exponential decay towards the steady-state conductance $g_{\infty}$ with a diffusion-like timescale $\tau = L^2/12D$, as in Ref. \cite{Rene_IonicNeuromorphic}. The equation for the conductance reads
\begin{equation}
\frac{\partial g\brac{\boldsymbol{\Delta}(t), t}}{\partial t} = \frac{g_{\infty}\brac{\boldsymbol{\Delta}(t)}-g\brac{\boldsymbol{\Delta}(t), t}}{\tau}
\label{eq:eqofmotconductance}
\end{equation}
with $\boldsymbol{\Delta}(t) = \brac{\Delta V\brac{t},\Delta P\brac{t}, \Delta \rho\brac{t}}$.
As in Ref. \cite{Rene_IonicNeuromorphic}, we consider the time scale relative to voltage-driven ion accumulation or depletion in the channel. We justify this choice by stating that pressure-driven dynamics is faster and recalling that we neglect chemically-driven transport. Due to the fact that the conductance depends on the history of the signals applied, this system is called memristor. Aside from the direct dependence of the conductance on external stimuli, there is also an implicit dependence on geometrical properties, such as length $L$ and base radius $R_b$, that we address in this work.

The PNN is composed of $\N$ nodes, an example is shown in \cref{fig:memristor_network}. 
\begin{figure}[h]
    \centering
    \includegraphics[width=0.8\columnwidth]{plots/memristor/memristor_network.pdf}
    \caption{An example of a network of five nodes connected by six memristors. The nodes are divided in input nodes (green circles), hidden nodes (grey nodes) and output nodes (purple circles) each of them is characterized by a triplet of physical quantities, representing the potential, pressure, and density. Th resulting stimuli on the memristor is given by the difference between base and tip signal, it thus depends on the orientation of the memristor. In the case of the memristor that connects the first and the second node, for example, the external stimuli are $\Delta V = V_2 - V_1$, $\Delta P = P_2 - P_1$ and $\Delta \rho = \rho_2 - \rho_1$. The zig-zag lines directed towards the input nodes indicate where the electrical signal is applied, while the zig-zag lines pointing outwards from the nodes indicate where the electrical signal is measured.}
    \label{fig:memristor_network}
\end{figure} 
Each node $\n\in[1,\N]$ is characterized by a value of electrical potential $V_{\n}$, pressure $P_{\n}$ and bulk density $\rho_{\n}$. Even though each node in the network is characterized by three physical properties, only electrical current flows through the network. This current propagates via the memristive elements, enabling modifications to the electric potential at other nodes. In contrast, pressure and density are externally imposed at each node and do not propagate through the network. The set of nodes is divided in three sets. A node belongs to the set of \textit{input nodes} if it is used to input current in the circuit, its value of voltage is known and fixed. After the inputs are imposed, the electrical signal travels through the network and the resulting output potential is read in the nodes belonging to \textit{output nodes}. The remaining nodes, that do not function as input or output, are included in the set of \textit{hidden nodes}.

The nodes in the PNN are connected by $\M$ memristive elements. Consequently, a memristor connecting two nodes experiences potential drop, pressure drop and concentration gradient given by the difference between the node where the base of the memristor is connected to, and the node to which the tip is linked. The memristor is characterized by its conductance state $g_{\m}$ with $\m\in[1,\M]$, consequence of both the geometry parameters and the external stimuli.

To solve numerically the resulting circuit, it is necessary to simultaneously address the circuit equations and the time evolution equation for the memristors' conductance. 
The circuit equations are handled using the open-source circuit solver \textsc{AHKAB}\cite{}, which computes the electric potentials at the nodes and the currents through the edges discretizing the differential equations in timesteps $\Delta t$. To incorporate the memristive elements, we approximate the time evolution of the conductances using a discretized form of \cref{eq:eqofmotconductance}, employing the Euler-forward method. Specifically, the conductance at a given time-step is calculated from its value at the previous time-step as:
\begin{multline}
g_{\m}\brac{\boldsymbol{\Delta}(t), t} = g_{\m}\brac{\boldsymbol{\Delta}(t), t-\Delta t} + \\
 \frac{g_{\m,\infty}\brac{\boldsymbol{\Delta}(t)-g_{\m}\brac{\boldsymbol{\Delta}, t}}}{\tau}\Delta t.
\end{multline}
In this work, we only apply constant voltage supplies to the input nodes of the circuit. Nonetheless, we notice that solving the dynamical equations is needed due to the dynamical change in the conductance of the memristors that, in turn, results in time-dependent potential drops across them. A more detailed illustration of the algorithm that solves the circuit, along with an analysis of the relaxation to the steady state of the memristor network, is provided in \cref{app:dyn_evol_steady_state}.

@@

We developed a training algorithm that makes use of the physical dependencies of memristor conductivity on both external stimuli and geometric properties to implement a physical neural network. These dependencies allow us to interpret such physical parameters as tuneable weights, analogous to standard artificial neural networks, and to encode learning directly into the physical structure of the system. As in traditional ANNs, the goal of the training algorithm is to modify these weights so that the network maps a given set of inputs to a corresponding set of desired outputs. In the case of the memristor network considered here, training aims to produce specific voltage values at designated output nodes, given fixed voltage inputs at the input nodes.
\begin{figure}[!h]
    \centering
    \includegraphics[width=0.8\columnwidth]{plots/training/two_types_weights.pdf}
    \caption{(a) Memristor network with weights defined in the nodes. (b) Memristor network where wegiths characterize each edge.}
    \label{fig:two_types_weights}
\end{figure} 
Crucially, the physical quantities that influence memristor conductance fall into two distinct categories: external stimuli, such as pressure or concentration, that are applied to the nodes, and geometric parameters, such as channel length or base radius, that are defined on the edges. As a result, we distinguish between two types of training strategies: \emph{node-defined weight training}, as illustrated in \cref{fig:two_types_weights}(a), and \emph{edge-defined weight training}, as shown in \cref{fig:two_types_weights}(b).

A modification to any of these weights affects the network's behaviour by altering the functional form of the memristor’s steady-state conductance, $g_{\infty}(\boldsymbol{\Delta})$, which depends on the local potential, pressure, and concentration drops across the device. Importantly, the \emph{entire functional shape} of $g_{\infty}$ plays a role in determining the network’s dynamics and ultimate steady state, as illustrated in \cref{app:dyn_evol_steady_state}.

In this work, we focus on training a single property as the weight, while keeping all other parameters fixed as initially defined. For example, we may select an external stimulus, such as the imposed pressure or the density at each node. Alternatively, we can choose the length or base radius of the channel. The training process is carried out by iterating a scheme, illustrated in \cref{fig:training_scheme}, which consists of two phases: a forward pass and a backward pass.
\begin{figure*}[t] 
    \centering
    \includegraphics[width=\textwidth]{plots/training/training_scheme.pdf}
    \caption{Scheme representing the two phases of the training algorithm, they are iteratively performed for a selected number of training steps.}
    \label{fig:training_scheme}
\end{figure*}
In the forward phase, the network is fed with specific electric potentials in the input nodes. Following both circuit laws and memristor's dynamics, the network reaches a steady state where electrical potentials can be measured in the output nodes. Clearly, the output of the network depends on the set of weights imposed on the network $\boldsymbol{w}$, regardless them being node or edge defined. The value of the outputs can be used to quantify the status of the network, meaning how much its answer differs from the desired one. Specifically, we define a cost function that computes the difference squared between the voltage in the output nodes and the values of desired voltages
\[
C\left( \boldsymbol{w} \right) = \sum_{i \in \Lambda} \brac{V_i\brac{\boldsymbol{w}} - V_i^D}^2,
\label{eq:cost_func}
\]
where $\Lambda$ represents the set of indices belonging to the output nodes, $V_i\brac{\boldsymbol{w}}$ the voltage measured on the output nodes and $V^D$ is a vector containing the values of desired output nodes.

The backward phase has the fundamental purpose of finding a set of weights that minimize the cost function. We follow the method of the \textit{steepest descent}, for which the update of the weights from one training step $s$ to the following $s+1$,
\begin{equation}
\boldsymbol{w}^{s+1} = \boldsymbol{w}^{s} - \alpha \boldsymbol{\nabla}C\brac{\boldsymbol{w}},
\label{eq:weight_update}
\end{equation}
results in a decrease in the cost function until a minimum is reached, for sufficiently small values of $\alpha$ \cite{steepest_descent_book}. The gradient in the update rule is computed using a \textit{finite differences method}
\begin{equation}
\boldsymbol{\nabla}C\brac{\boldsymbol{w}} = \frac{C\brac{\boldsymbol{w}+\boldsymbol{\Delta w}}-C\brac{\boldsymbol{w}}}{\Delta w},
\label{eq:gradient_cost_func}
\end{equation}
where $\boldsymbol{\Delta w}$ is a vector with the same number of entries as the weight vector, but every entry has a value $\Delta w$.

In the training algorithm, we first initialize the system with the memristor parameters specified in \cref{sec:memristor_network}, an initial pressure of $1bar$ and concentration of $0.2mM$ at each node. With the knowledge of input and desired output voltage values, it is possible to perform the first forward pass, from which we obtain the cost function $C\brac{\boldsymbol{w}}$, that we store since it is needed in the computation its gradient. Subsequently, we proceed with the backward pass. We compute each entry of the gradient in \cref{eq:gradient_cost_func} by selecting the corresponding weight, a node or an edge parameter, increment its value of $\Delta w$ and applying a forward step to obtain the corresponding entry of $C\brac{\boldsymbol{w}+\boldsymbol{\Delta w}}$. With the stored value of $C\brac{\boldsymbol{w}}$, we are able to compute the gradient in \cref{eq:gradient_cost_func} and use it to calculate the weight update in \cref{eq:weight_update}.

Although performed numerically in this work, we highlight that the feedforward pass is a physical process in which the system naturally relaxes to its steady state. Moreover, also the backward pass we use can be translated to a physical event since involves performing the forward pass a number of time as the number of the weights. Once the gradient of the cost function is determined and stored, only the simple algebraic calculations of \cref{eq:weight_update} are left to be computed. We propose this method as a hybrid physical and in-silico training algorithm, as a future experimental approach to the problem.

@@

We initialize the system with pressures $P_1 =  P_2 = 1 bar$ and ion concentration $\rho_1 = \rho_2 = 0.2 mM$ on each node, length $L_1 = L_2 = 10\mu m$ and base radius  $R_{b1} = R_{b2} = 200 nm$ on each edge. After choosing one of these four physical quantities as training parameter $\boldsymbol{w} = \brac{w_1, w_2}$, we implement the training as described in \cref{sec:training} that modifies them, while the other parameters are kept constant. 
In \cref{fig:mse_weights_vd} we show that the cost function normalized to its initial value decreases of many order of magnitude during the training for each of the chosen weight. The values of the chosen set of weights simultaneously changes from the initial condition, as shown in Figures [], adjusting the behaviour of the network to the desired one. If the values of a chosen type of weight at the end of the training are imposed to the network together with the voltage inputs, the network will autonomously relax to a steady state that coincides with the desired state. To exemplify this point, the inset in Figure shows the physical relaxation of the voltage divider when the trained values of pressure, which correspond to the three values in the last training step, are imposed on the nodes together with the constant inputs voltages. The inset shows that the output voltage $V_2$ autonomously relaxes to the desired value of $V_D=4V$ in a time scale characteristic of the memristor.

\begin{figure}[h]
    \centering
    \includegraphics[width=\columnwidth]{plots/voltage_divider/mse_weights_vd.pdf}
    \caption{(a) Behaviour of the cost function during the training of the voltage divider. Different colors indicate that the training is done with a specific choice of training weights, as indicated in the legend. (b) Pressures imposed on the three nodes of the voltage divider during training. The inset shows the relaxation to the steady state of the output voltage $V_2$ to the desired value $V_D$ when the pressures imposed are the values obtained in the last training step, and the input nodes are stimulated with fixed voltage. (c) Ion concentration densities imposed on the three nodes of the voltage divider during training. (d) Length of the memristors during training. (e) Base radius of the memristors during training.}\label{fig:mse_weights_vd}
\end{figure}

In order to investigate the adaptability of the algorithm, we trained the system with a set of three consequent desired voltages $V_D=\{1V, 3V, 4V\}$. In \cref{fig:evolution_targets} we show how the output voltage $V_2$ subsequently adapts to the desired voltages during the training of the pressures in the nodes, the evolution of which can be seen in \cref{fig:evolution_targets}. It is evident that the system can be retrained at any stage to a new desired objective making it reusable. Any other weight type can be chosen, depending on the necessity.

\begin{figure}[h]
    \centering
    \includegraphics[width=0.8\columnwidth]{plots/voltage_divider/evolution_targets.pdf}
    \caption{(a) The training of a memristor voltage divider to a set of desired voltage outputs $V_D = {1V, 3V, 4V}$, indicated by dashed light purple lines. The output voltage $V_2$, indicated by dark purple dots, adapts to the desired output during the training. The input voltages $V_1$ and $V_3$ are constant at $5V$ and $0V$ respectively. (b) The dots with different shades of grey show the values of pressures, chosen as weights, imposed on the three nodes during training.}\label{fig:evolution_targets}
\end{figure}

@@

To generalize the training objective, we consider a memristor network constructed with a more complex geometry. \cref{fig:mse_general}(a) illustrates a network with three input nodes and two output nodes. The network is trained to produce output voltages of $[3,V, 4,V]$ when provided with input voltages of $[0,V, 2,V, 5,V]$.  Reaching a desired input-output voltage mapping in a generic electric circuit can be useful for various electronic purposes \cite{voltage_controlled_oscillator, onchip_powerdistributed}.

\begin{figure}[h]
    \centering
    \includegraphics[width=0.5\textwidth]{plots/general_network/mse_general.pdf}
    \caption{(a) Generic network with $9$ nodes and $15$ edges with inputs $[V_{2}, V_{4}, V_{9}] = [0V, 2V, 5V]$ and desired outputs $[V_{5}^D, V_{7}^D] = [3V, 4V]$, delta weights a third of their initial value and learning rates $[0.001, 1, 0.0001, 0.001]$. (b) Cost function during training of generic graph for input-output mapping task. The different colors indicate different weight choice.}\label{fig:mse_general}
\end{figure}

The result of the training is presented in \cref{fig:mse_general}(b), which shows how the cost function responds to the weight updates performed in the algorithm, for each of the four different weights. All of the possible different choices of weight result in a trained set that decreases the initial error of many orders of magnitudes.
